On Principles for Government Open Data Curation

### 2023-06-10 talk/slides.Rmd
minor edits

- Open Data is viewed as a right for all [@janssen2012benefits; @wang2016adoption].
- Cities lead open data efforts wtih >50 cities with formal programs.
- NYC is a forerunner with Open Data Law in 2012 [@zuiderwijk2014open].
- Today NYC Open Data has 3511 datasets from 98 Agencies, accesible online.
- Used in many research
    + Health [@cantor2018facets, shankar2021data]
    + Smart city [@neves2020impacts]
    + Transportation [@gerte2019understanding]

@@

- 311 Service Request data is the most viewed dataset at NYC Open Data.
- Used in STAT 3255/5255 Introduction to Data Science at UConn
- In this study, we exported 1-31 March 2023 311 SRs for analysis (~251K records).
- Data curation is imporant issue for open data such as the 311 SRs [@witt2009constructing; @borgman2012conundrum; @mclure2014data; @barns2016mine; @hart2016ten; @beheshti2019datasynapse; @shankar2021data].

@@

## Invalid Data: Zipcodes (One of the last 5 columns)
```{r zip2}
badZipcode  <- subset(d311, ! zip_codes %in% validZipcodes,
                      select = c("zip_codes", "agency"))
nrow(badZipcode)  # mostly invalid!
```

@@

## Invalid Data: Open Data Channel
```{r channel}
validChannel <- c("MOBILE", "ONLINE", "PHONE", "OTHER")
badChannel_type <- subset(d311, ! open_data_channel_type %in% validChannel,
                          select = c("open_data_channel_type", "agency"))
head(table(badChannel_type$agency)) ## UNKNOWN
```

@@

@@

+ Improve dictionary
    - Code categorical variables (e.g., agency) 
	- Code descriptive fields (e.g., complain type; vehicle type)
    - Make the descriptions informative (e.g., zip code)
+ Impose validity check
    - date (logical check)
    - legitimate values 
	- descriptive fields (drop-down menu, possibly multi-step)
    - missing value: coding; validity
+ Improve data storage/transportation efficiency
    - Eliminate redundant columns
    - Reduce storage with no information loss
+ Feedback: 
    - Open review for open data
	- Interaction with the maintenance team

>>>>

- Open Data is viewed as a right for all [@janssen2012benefits; @wang2016adoption].
- Cities lead open data efforts wtih >50 cities with formal programs.
- NYC is a forerunner with Open Data Law in 2012 [@zuiderwijk2014open].
- Today NYC Open Data has 3511 datasets from 98 Agencies, accesible online 
  (\url{https://opendata.cityofnewyork.us}).
- Used by many researchers:
    + Health [@cantor2018facets, shankar2021data]
    + Smart city [@neves2020impacts]
    + Transportation [@gerte2019understanding]

@@

- 311 Service Request data is the most viewed dataset at NYC Open Data.
- NYC 311 is a unified frontline for all service requests to the NYC government agents.
- The data was used in STAT 3255/5255 Introduction to Data Science at UConn, with student presentations at 2022/2023 NYC Open Data Week.
- In this study, we exported 1-31 March 2023 311 SRs for analysis (~251K records).
- Data curation is critical for open data such as the 311 SRs [@witt2009constructing; @borgman2012conundrum; @mclure2014data; @barns2016mine; @hart2016ten; @beheshti2019datasynapse; @shankar2021data].
- Issues
    - Validity
	- Efficiency

@@

## Invalid Data: Zipcodes (One of the last 5 columns)
```{r zip2}
badZipcode  <- subset(d311, ! zip_codes %in% validZipcodes,
                      select = c("zip_codes", "agency"))
nrow(badZipcode)  # mostly invalid! One possibility: debt collectors
```

@@

## Invalid Data: Open Data Channel
"Due to pandemic call handling modifications, the ‘Open Data Channel Type’ values may not accurately indicate the channel the Service Request was submitted in for the period starting March 2020."

```{r channel}
validChannel <- c("MOBILE", "ONLINE", "PHONE", "OTHER")
badChannel_type <- subset(d311, ! open_data_channel_type %in% validChannel,
                          select = c("open_data_channel_type", "agency"))
head(table(badChannel_type$agency)) ## UNKNOWN
```

@@

## Redundant Data: Agence, Agency Name, and Other Categorical Variables
```{r agent}
head(d311[, c("agency", "agency_name")])
head(sort(table(d311$complaint_type), decreasing = TRUE)) # 174 types
```

@@

+ Improve dictionary
    - Code categorical variables (e.g., agency) 
	- Code descriptive fields (e.g., complain type; vehicle type)
    - Make the descriptions informative (e.g., zip code)
+ Impose validity check
    - date (logical check)
    - legitimate values 
	- descriptive fields (drop-down menu, possibly multi-step)
    - missing value: coding; validity
+ Improve data storage/transportation efficiency
    - Eliminate redundant columns
    - Reduce storage with no information loss

## Outlook

- Open review for open data

- Interaction with the maintenance team

- Open data curation protocol

- Open data analytic portal
